---
title: "SE of SE(X-bar) - follow-up 1"
author: 'Marta Karas'
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Methods: se of se(X-bar)

We agree that

$$
\begin{aligned}
\hat{S E}\left(\bar{X}_{1}\right) & =\hat{s}_{1} / \sqrt{n_{1}} \\
\operatorname{Var}\left[\hat{S E}\left(\bar{X}_{1}\right)\right] & =\operatorname{Var}\left[\hat{s}_{1} / \sqrt{n_{1}}\right]=\operatorname{Var}\left(\hat{s}_{1}\right) / n_{1}
\end{aligned}
$$

#### Approximate solution (Marta's attempt)

Assuming $X_i \sim N(\mu, \sigma^2)$, we arrived at 

$$
\begin{aligned}
\operatorname{Var}\left(\hat{s}_{1}\right)
&  \approx \sqrt{Var(s_1^2)} \\
\operatorname{Var}\left[\hat{S E}\left(\bar{X}_{1}\right)\right]
&  \approx  \left( \sqrt{\operatorname{Var}\left(\hat{s}_{1}^{2}\right)} \right) / n_1
= \sqrt{\frac{2 \sigma^{4}}{\left(n_1-1\right) n_1^{2}}} \\
& = \frac{\sigma^2}{n_1} \cdot \sqrt{\frac{2}{\left(n_1-1\right) }}% \\
%& =  \frac{\sigma^2}{n_1} \cdot c_{4}(n_1) \cdot \frac{\Gamma((n_1-1) / 2)}{\Gamma(n_1 / 2)}
\end{aligned}
$$

#### Exact solution (Lacey's attempt)

Below we will use notation

$$
\begin{aligned}
c_{4} & = \sqrt{\frac{2}{n-1}} \cdot  \frac{\Gamma\left(\frac{n}{2}\right)}{\Gamma\left(\frac{n-1}{2}\right)} = \frac{E(\chi_{n_1-1})}{\sqrt{n-1}} \\
\Rightarrow E(\chi_{n_1-1}) & = c_{4} \cdot \sqrt{n-1} \\
\Rightarrow \left( E(\chi_{n_1-1}) \right)^2 & = c_{4}^2 \cdot (n-1).
\end{aligned}
$$

Assuming $X_i \sim N(\mu, \sigma^2)$, we arrived at 

$$
\begin{aligned}
\operatorname{Var}\left(\hat{s}_{1}\right) / n_{1}
& = \operatorname{Var}\left(  \frac{\chi_{n_1-1} \sigma}{\sqrt{n_1-1}} \right)  / n_{1} \\
& = \frac{\sigma^2}{n_1(n_1-1)} \cdot Var(\chi_{n_1-1}) \\
& = \frac{\sigma^2}{n_1(n_1-1)} \cdot \left[(n_1-1) - \left( E(\chi_{n_1-1}) \right)^2 \right] \\
& = \frac{\sigma^2}{n_1(n_1-1)} \cdot \left[(n_1-1) - c_{4}^2 \cdot (n-1) \right] \\
& = \frac{\sigma^2}{n_1} \cdot \left(1 - c_{4}^2 \right)
\end{aligned}
$$

Note the above aligns with [wiki article *Unbiased estimation of standard deviation*](https://en.wikipedia.org/wiki/Unbiased_estimation_of_standard_deviation) which states

$$
\begin{aligned}
 sd(\hat{s_1}) = \sigma\sqrt{1 - c^2_4}
\end{aligned}
$$

#### The difference between the two approaches

We see that the two solutions are off
$$
\begin{aligned}
\frac{\sigma^{2}}{n_{1}} \cdot \sqrt{\frac{2}{\left(n_{1}-1\right)}} \neq \frac{\sigma^{2}}{n_{1}} \cdot\left(1-c_{4}^{2}\right)
\end{aligned}
$$

Let's check what are the numerical values of those scalars.


## Checks with code 

* Note, that for large $n$, function `gamma` is not executable, hence computing $var(\chi_{n_1-1})$ is not executable

```{r}
options(scipen=999)

chi_var <- function(k){
  k - (sqrt(2) * gamma((k+1)/2) / gamma(k/2))^2
}

n_grid <- seq(from = 100, to = 1000, by = 100)
unlist(gamma(n_grid/2))
chi_var(n_grid)
```

* There is [c4](https://www.rdocumentation.org/packages/IQCC/versions/1.0/topics/c4) function tho, which is also not executable

```{r, message=FALSE}
# install.packages("IQCC")
library(IQCC)
c4(n_grid)
```

* Compare the scalars from the above 

```{r}
scalar_1 <- function(n){
  sqrt(2/(n-1))
}

scalar_2 <- function(n){
  1 - (c4(n))^2
}
n_grid <- seq(20, 200, by = 20)
```

```{r}
round(scalar_1(n_grid), 5)
round(scalar_2(n_grid), 5)
```

## Methods: se and var of se(X)

Assuming $X_i \sim N(\mu, \sigma^2)$, we can pull from the above

$$
\begin{aligned}
\operatorname{Var}\left(\hat{s}_{1}\right) 
& = \sigma^2 \cdot \left(1 - c_{4}^2 \right), \\
\operatorname{sd}\left(\hat{s}_{1}\right) 
& = \sigma \cdot \sqrt{1 - c_{4}^2}.
\end{aligned}
$$

where $c_4(n)$ can be approximated as 

$$
\begin{aligned}
c_{4}(n)=\sqrt{\frac{2}{n-1} \frac{\Gamma\left(\frac{n}{2}\right)}{\Gamma\left(\frac{n-1}{2}\right)}}=1-\frac{1}{4 n}-\frac{7}{32 n^{2}}-\frac{19}{128 n^{3}}+O\left(n^{-4}\right)
\end{aligned}
$$

